 ---
title: 'CSCI E-106:Assignment 7'
output:
  pdf_document: default
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Due Date: November 9, 2020 at 7:20 pm EST
### Instructions

Students should submit their reports on Canvas. The report needs to clearly state what question is being solved, step-by-step walk-through solutions, and final answers clearly indicated. Please solve by hand where appropriate.

Please submit two files: (1) a R Markdown file (.Rmd extension) and (2) a PDF document, word, or html generated using knitr for the .Rmd file submitted in (1) where appropriate. Please, use RStudio Cloud for your solutions.

----------------------------------------------------


## Problem 1

Use the fortune data under the faraway r library, data(fortune,package="faraway"). The wealth in billions of dollars for 232 billionaires is given in fortune. (50 points, 10 points each) (Hint: refer to the interaction.pdf and rmd files for details)

a-)Plot the wealth as a function of age using a different plotting symbol for the different regions of the world.

```{r}
rm(list = ls())

setwd("~/OneDrive/courses/e106/e106_hes2020/HW/HW7") # change to local path before running

library(ggplot2)
library(MASS)
library(onewaytests)
library(lmtest)

data("fortune", package = 'faraway')
head(fortune)
summary(fortune) # shows 5 NA values 
str(fortune)

# remove NA values
fortune <- na.omit(fortune)
summary(fortune) # na values gone

plot(wealth~age, fortune, pch = unclass(region), main = 'Wealth vs Age')

```

b-)Plot the wealth as a function of age with a separate panel for each region.

```{r}
ggplot(aes(x=age, y = wealth), data = fortune) + geom_point() + facet_wrap(~region)
```

c-)Determine a transformation on the response to facilitate linear modeling.

```{r}
reg1c <- lm(wealth~., data = fortune)
summary(reg1c)
boxcox(reg1c, lambda = seq(-2, 2, by = 0.1))

```

lambda has MLE value at -1 as seen above. This means that we take reciprocal of y

d-)What is the relationship of age and region to wealth?

Age and wealth appear to have a linear relationship though in all cases the values are concentrated between 40 and 80 years with a few outliers present.

Regions E & U have high concentration of wealthy people while region M has the lowest. Region M also has the largest outlier value.

Across all regions the relationship appears linear with concentration around the 40 to 80 age bracket

e-)Check the assumptions of your model using appropriate diagnostics.

```{r}
# Check for assumptions of simple regression model

# SW Test for normality
# Ho : Errors are normally distributed
# Ha : Errors are not normally distributed
shapiro.test(reg1c$residuals)

# BP Test for constancy of error variance
bptest(reg1c)
# Ho : Errors are normally distributed; no heteroskedasticicty
# Ha : Errors are not normally distributed
# Decision Rule: p-value > 0.05 ==> Accept Ho

# Residual plots
newer.par = par(mfrow = c(2,2))
plot(reg1c)
newer.par

```




```{r}
# Fit transformation using reciprocal as per BoxCox test
reg1e <- lm(1/wealth~., data = fortune)
summary(reg1e)

# Check for assumptions of simple regression model

# SW Test for normality
# Ho : Errors are normally distributed
# Ha : Errors are not normally distributed
shapiro.test(reg1e$residuals)

# BP Test for constancy of error variance
bptest(reg1e)
# Ho : Errors are normally distributed; no heteroskedasticicty
# Ha : Errors are not normally distributed
# Decision Rule: p-value > 0.05 ==> Accept Ho

# Residual plots
newer.par = par(mfrow = c(2,2))
plot(reg1e)
newer.par
```

## Problem 2

Refer to the CDI data set. A regression model relating serious crime rate (Y, total serious crimes divided by total population) to population density ($X_1$, total population divided by land area) and unemployment rate ($X_3$) is to be constructed. (30 points, 10 points each)

a-) Fit second-order regression model. Plot the residuals against the fitted values. How
well does the second-order model appear to fit the data? What is $R^2$?

```{r}
cdi <- read.csv('CDI Data.csv')
head(cdi)
summary(cdi)
str(cdi)

crime_rate <- cdi$Total.serious.crimes/cdi$Total.population
pop_den <- cdi$Total.population/cdi$Land.area
unemp <- cdi$Percent.unemployment

# fit 1st order model
reg2a_1 <- lm(y~x1+x2)
summary(reg2a_1)

# fit the 2nd order model
y <- crime_rate
x1 <- scale(pop_den, scale = FALSE)
x3 <- scale(unemp, scale = FALSE)
x11 <- x1^2
x32 <- x3^2
x13 <- x1*x3

reg2a_2 <- lm(y~x1+I(x1^2)+x3+I(x3^2)+(x1*x3))
summary(reg2a_2)

```

b-) Test whether or not all quadratic and interaction terms can be dropped from the regression
model; use $\alpha$ = .01. State the alternatives, decision rule, and conclusion.

```{r}
anova(reg2a_1, reg2a_2)
```
c-) Instead of the predictor variable population density, total population ($X_1$ and land area
($X_2$) are to be employed as separate predictor variables, in addition to unemployment rate ($X_3$). The regression model should contain linear and quadratic terms for total population, and linear terms only for land area and unemployment rate. (No interaction terms are to be included in this model.) Fit this regression model and obtain $R^2$. Is this coefficient of multiple determination substantially different from the one for the regression model in part (a)?

```{r}
crime_rate <- cdi$Total.serious.crimes/cdi$Total.population
populaiton <- cdi$Total.population
land_area <- cdi$Land.area
unemp <- cdi$Percent.unemployment

y <- crime_rate
x1 <- scale(populaiton, scale = FALSE)
x2 <- scale(land_area, scale = FALSE)
x3 <- scale(unemp, scale = FALSE)

x12 <- x1

reg2c <- lm(y~x1 + I(x1^2)+x2+x3)
summary(reg2c)
```

## Problem 3 

Refer to the CDI data set. The number of active physicians (Y) is to be regressed against total population ($X_1$), total personal income ($X_2$), and geographic region ($X_3$, $X_4$ , $X_5$). (20 points, 10 points each)

("("Geographic region classification is that used by the U.S. Bureau of the Census, where: 1 = NE, 2 = NC, S = 3, W=4")

a-)  Fit a first-order regression model. Let $X_3$ = 1 if NE and 0 otherwise, $X_4$ = 1 if NC and 0 otherwise, and $X_5$ = 1 if S and 0 otherwise. Examine whether the effect for the northeastern region on number of active physicians differs from the effect for the north central region by constructing an appropriate 90 percent
confidence interval. Interpret your interval estimate.

```{r}
y <- cdi$Number.of.active.physicians
x1 <- cdi$Total.population
x2 <- cdi$Total.personal.income
x3 <- as.numeric(cdi$Geographic.region==1)
x4 <- as.numeric(cdi$Geographic.region==2)
x5 <- as.numeric(cdi$Geographic.region==3)

reg3a <- lm(y~x1+x2+x3+x4+x5)
summary(reg3a)

```

b-) Test whether any geographic effects are present; use $\alpha$ = .10. State the alternatives, decision
rule, and conclusion. What is the P-value of the test?

```{r}
reg3b <- lm(y~x1+x+2)
summary(reg3b)

anova(reg3b, reg3a)
```