---
title: "CSCI_E_106 Section 2"
author: "Rafael GT"
date: "9/17/2020"
output: html_document
---
## Before you start:

*Create a sub-directory in your computer
*Open CANVAS and look for the data 
*Read specs see below-rephrase the problem with your own words
*Master story telling and enjoy the process

# Review from Class 1:
1. Toluca Example (5 minutes)

The Toluca company manufactures refrigerator equipment and parts in lots of
various sizes. 

Company officials want the size of the lots to be optimized
subjet to their production cost.

The key driver of costs is the number of labor hours to produce the lots.

Management collects data on the size of 25 lots and corresponding hours.

Managers believe the information behaves like normal.
 
Step 1.- State the equation for the normal error model for the problem.
Assumptions:
\[Y_i=\beta_o+\beta_iX_i+\varepsilon_i \]

\[
  \varepsilon_i\sim iid \mathcal{N}(\mu,\,\sigma^{2})\,.
\] 
## R Studio Cloud  -R Commands
Step 2.- Set your R-cloud environment:
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(readr)
rm(list=ls())
```
Step3.- Upload Toluca Data to R Studio Cloud:
```{r}
toluca_data <- read_csv("toluca_data.csv")
head(toluca_data)
View(toluca_data)
attach(toluca_data)
```
Step4 fitting the regression model: 
```{r}
toluca.reg <- lm(workhrs ~ lotsize)
```
Step5 getting the summary regression output:
```{r}
summary(toluca.reg)
```
Step6 getting the Analysis of Variance table (ANOVA):
```{r}
anova(toluca.reg)
```
Step7 getting the fitted values:
```{r}
fitted(toluca.reg)

```
Step8 getting the residual values (why not the errors?):
```{r}
names(toluca.reg)
resid(toluca.reg)
toluca.reg$residuals
```
Step9 overlaying the regression line on the scatter plot of X and Y:
```{r}
plot(lotsize, workhrs)
abline(toluca.reg)
```
```
The error (or disturbance) of an observed value is the deviation of the observed
value from the (unobservable) true value of a quantity of interest (for example,
a population mean), and the residual of an observed value is the difference 
between the observed value and the estimated value of the quantity of interest.

2. Beer Example (5 minutes)

The Beer manufacturing industry is very competitive. 

When a beer stays too long in storage its flavor rapidly decays ruining
the consumer experience.

The main drivers for the duration of the good taste of a beer are the beer’s 
antioxidant compounds.

In general beers with high “Total Phenolic Content (TPC)” taste good.

Management collects relevant data on 40 lager beers, and they believe the
information behaves like normal.

The objective is to find the best beer’s antioxidant compounds.
(Perform one factor analysis by finding the best variable to explain Total
Phenolic Content (TPC)). 


*Repeat Steps 1 and 2 above.
Step3.- Upload Beer Data to R Studio Cloud:
```{r}
Beerdata<- read_csv("lager_antioxidant.csv")
head(Beerdata)
View(Beerdata)
attach(Beerdata)
Beerdata=data.frame(Beerdata)
```
Step4.-  Fit one variable regression model with TPC as a dependent variable
against remaining variables above, as an independent variable one at time
For example, TPC = bo + b1 MA, TPC=b0+b1 DSA and so on:
```{r}
f1<-lm(tpc~ma,data=Beerdata)
f2<-lm(tpc~dsa,data=Beerdata)
f3<-lm(tpc~asa,data=Beerdata)
f4<-lm(tpc~orac,data=Beerdata)
f5<-lm(tpc~rp,data=Beerdata)
f6<-lm(tpc~mca,data=Beerdata)
```
Step5.-Let know the computer how to organize the information.
Use an R-function. The "par(mfrow)" function is handy for creating
a simple multi-paneled plot:
```{r}
par(mfrow=c(2,3))
plot(ma,tpc);abline(f1)
plot(dsa,tpc);abline(f2)
plot(asa,tpc);abline(f3)
plot(orac,tpc);abline(f4)
plot(rp,tpc);abline(f5)
plot(mca,tpc);abline(f6)
```
Step6.-Use an R-function to find the best fit.
Each function needs to input the parameters "function(here the variables)", 
to know where the results are going to be saved ("data.framework"), when
to stop (for i...). What is the equation or operations you want to do:
```{r}

prg1<-function(x){         
  out1<-list({})         
  out2<-data.frame(matrix(0,nrow=6,ncol=2)) 

           for (i in 1:6){                   

                    out1[[i]]<-lm(x[,2]~x[,2+i],data=x)                   

                    out2[i,]<-cbind(dimnames(x)[[2]][2+i],round(sum(out1[[i]]$residuals^2),2)) 

                              }     
   dimnames(out2)[[2]]<-c("X","SSE")    

   list(out1,out2)
}
prg1(Beerdata)

```