---
title: 'Practice Final Exam'
output:
  pdf_document: default
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Instructions

Open book and open notes exam ( textbooks (print or pdf), lecture slides, notes, practice exam, homework solutions, and TA slides, including all Rmd's*).

You are allowed to use RStudio Cloud (https://rstudio.cloud.) , Microsoft Word, Power Point and PDF reader, and canvas on your laptop.

Proctorio is required to start this exam. If you are prompted for an access code, you must Configure Proctorio  on your machine. 

Please read the list of recording and restrictions provided by Proctorio carefully before taking the exam 

Please pay attention any timing and technical warnings that popped up your screen 

The exam will be available from Monday December 14th at 8 am EST through Tuesday December 15th at 8:00pm EST.
Once you start the exam, you have to complete the exam in 3 hours or by Tuesday December 15th at 8:00pm EST, whichever comes first.

In order to receive full credit, please provide full explanations and calculations for each questions 

Make sure that you are familiar with the procedures for troubleshooting exam issues Preview the document
Make sure you submit both .Rmd and (knitted) pdf or html files.


You need to have a camera on your laptop.

----------------------------------------------------

```{r}
rm(list = ls())

setwd("/cloud/project/HW/practice_final")

library(olsrr) # regression residuals
library(datasets) # datsets
library(leaps) # regsubsets
library(car) # avplot
library(ggplot2) # graphics
library(caret) # cross validation
library(onewaytests) # Regression Diagnostics
library(lmtest) # Regression Diagnostics
library(QuantPsyc) # #standardized regression
library(faraway) # datasets
library(MASS) # negative binomial; robust regression
library(ResourceSelection) # goodness of fit test
library(glmnet) # ridge, lasso, ElasticNet, Logistic, Poisson, Cox
library(olsrr) # regression tests
library(lmtest) # time series
library(Hmisc) # Lag
library(orcutt) # cochrante-orcutt procedure
library(HoRM) # hildreth Lu
library(C50) # decision tree
library(partykit) # better visualization
library(gmodels) # cross table
library(randomForest) #RF
library(vcd) # confusion matrix
library(rpart) # regression tree
library(rpart.plot) # plots of decision trees
library(neuralnet) # neural net
library(ResourceSelection) # goodness of fit test

```

## Problem 1

Use the question1 data, fit the regression model on Y by using all the variables (X6 and X7 are categorical variables). Create development sample (70% of the data) and hold-out sample (30% of the data).  Perform statistical tests, use graphs or calculate the measures (e.g. VIF, Leverage Points, Cook’s Distance) for questions below. Use the development sample for part a to d. Use the hold-out sample for part e.
  
a-) Is the model significant? Is there a Multicollinearity in the data? Are the errors Normally distributed with constant variance?

```{r}
# load data, examine columns and check for NA, etc.
p1 <- read.csv('Practice Final Question 1.csv')
head(p1)
dim(p1)
str(p1)
summary(p1)

# create train and test set
set.seed(1023)
ind <- sample(1:nrow(p1), round(0.7*nrow(p1)))
tr_p1 <- p1[ind,]
tes_p1 <- p1[-ind,]
#check
dim(tr_p1)

# fit regression
reg1a <- lm(Y~., data = tr_p1)
summary(reg1a)

# plot residuals
newer.par <- par(mfrow = c(2,2))
plot(reg1a)
newer.par

vif(reg1a)
```

From above we see that the model has F-stat = 11.04 with p-value = 6.65e-11 which means that the overall model is significant

Individual coefficients x3, x6 and x10 are not signifiant.

There is evidence of multicollinearity for x5, x8 and x9; we use lasso since some variables are not significant and we have evidence of multicollinearity

```{r}
x <- model.matrix(Y~., data = tr_p1) [,-c(1)]; x[1:5,]
y <- tr_p1$Y

lasso_1a <- glmnet(x, y, alpha = 1, nlambda = 100, lambda.min.ratio = 0.0001)
plot(lasso_1a, xvar = 'norm', label = TRUE)

# cross validation
cv_lasso_mod <- cv.glmnet(x, y, alpha = 1, nlambda = 100, lambda.min.ratio = 0.0001)
plot(cv_lasso_mod)

best_lambda_lasso <- cv_lasso_mod$lambda.min
best_lambda_lasso

# 2 ways to predict coefficients
coef(cv_lasso_mod, s = 'lambda.min')
```

```{r}
reg1a_best <- ols_step_best_subset(reg1a)
plot(reg1a_best)

which.max(reg1a_best$adjr); which.min(reg1a_best$cp); which.min(reg1a_best$aic)
```

We choose model 7 for the final model; create df and perform regression below

```{r}
min_aic <- which.min(reg1a_best$aic)
tr1a_aic_cols <- unlist(strsplit(reg1a_best$predictors[min_aic], split = ' ')); tr1a_aic_cols

tr1a_aic_df <- tr_p1[tr1a_aic_cols]
Y <- tr_p1$Y
tr1a_aic_df <- cbind(Y, tr1a_aic_df); head(tr1a_aic_df)

reg1a_final <- lm(Y~., data = tr1a_aic_df)
summary(reg1a_final)

# plot residuals
newer.par <- par(mfrow = c(2,2))
plot(reg1a_final)
newer.par

vif(reg1a_final)

```
```{r}
par(mfrow = c(1,1))
boxcox(reg1a_final, lambda = seq(-2,2,0.1))
```
Take reciprocal as per boxcox

```{r}
reg1a_final2 <- lm(1/Y~., data = tr1a_aic_df)
summary(reg1a_final2)

# plot residuals
newer.par <- par(mfrow = c(2,2))
plot(reg1a_final)
newer.par

vif(reg1a_final2)
```


Residuals vs fitted values chart seems to be normally distributed while QQ plot chart is not at 45 degrees. There is also evidence of some outliers.

Hence regression assumptions relating to normal distribution of error terms and model fitting all but 1 or few outliers are not satisfied.

b-) Are there any influential or outlier observations?

```{r}
influence.measures(reg1a_final2)
ols_plot_dffits(reg1a_final2)
ols_plot_cooksd_chart(reg1a_final2)
```

From above charts we see that points 69, 9 are potential outliers.

c-) Can X5, X6, and X7 be dropped from the model? Perform the statistical test and state your final model.

```{r}
# reduced model
reg1c <- lm(Y ~ X1 + X2 + X3 + X4 + X8 + X9 + X10, data = tr_p1)
# summary(reg1c)

summary(reg1a)$r.squared; summary(reg1c)$r.squared
# anova test to check if coefficients can be dropped
anova(reg1c, reg1a)

```
Ho : x5, x6, x7 = 0; x5, x6 and x7 can be dropped
Ha : x5, x6, x7 != 0; x5, x6 and x7 cannot be dropped

from above we see that p-value = 0.003 which is less than alpha = 0.05 (assumed). Hence we fail to reject Ho and conclude that these terms need to be retained.

d-) Develop an alternative model by using the Regression Tree and compare the performance against the regression model built in part a. 

```{r}
tree1d <- rpart(Y~., data = tr_p1)
plot_tree1d <- rpart.plot(tree1d, digits = 5, fallen.leaves = TRUE, type = 2, extra = 101)
```

f-) Score the model on hold-out sample, and re-calibrate the model on the holdout sample. Compare the results against the final model derived part c.


## Problem 2

Use the question2 data set to answer this question. We are interested in predicting  (Y) the number of customers who complained about the service.
 
a-) Build a model to predict the number of complaints, perform the statistical tests that shows that model is significant

```{r}
# load data, examine columns and check for NA, etc.
p2 <- read.csv('Practice Final Question 2.csv')
head(p2)
dim(p2)
str(p2)
summary(p2)

reg2a_lm <- lm(Y~., data = p2)
summary(reg2a_lm)

# plot residuals
newer.par <- par(mfrow = c(2,2))
plot(reg2a_lm)
newer.par

vif(reg2a_lm)
```

b-)   Find the predicted number complaints given the independent variables below and predict 95% confidence interval

X1=606 X2=41393	X3=3	X4=3.04	X5=6.32

```{r}
test2a <- data.frame(X1=606, X2=41393, X3=3,	X4=3.04,	X5=6.32); test2a
predict(reg2a_lm, test2a, interval = 'confidence', level = 0.95)
```


## Problem 3 

Use question 3 data sets. Monthly data on amount of billings (Y) and on number of hours of staff time (X) for the 20 most recent months are recorded.

a-) Build a model to predict Y based on the independent variables and test if there is an autocorrelation persists in the data. If autocorrelation persists, remediate the autocorrelation. 

```{r}
# load data, examine columns and check for NA, etc.
p3 <- read.csv('Practice Final Question 3.csv')
head(p3)
dim(p3)
str(p3)
summary(p3)

# durbin-watson test for autocorrelation
dwtest(Y~X, data = p3)

```
Ho : No autocorrelation
Ha : Autocorrelation > 0

p-value = 0.002891 which is less than 0.05; hence we reject Ho. This implies there is autocorrelation.

Use cochrane orcutt procedure to remedy autocorrelation

```{r}
# Prediction
f <- lm(Y~X, data = p3)
et <- f$residuals
et1 <- Lag(et, shift = 1)

d1 <- sum(na.omit(et1*et))
d2 <- sum(na.omit(et1)^2)

rho <- d1/d2
rho

ytnew <- p3$Y - rho*Lag(p3$Y, shift = 1)
xtnew <- p3$X - rho*Lag(p3$X, shift = 1)

f1 <- lm(ytnew~xtnew)
summary(f1)
mse <- summary(f1)$sigma^2; mse

dwtest(f1)
```

b-) X (Staff time) in month 21 is expected to be 3.625 thousand hours. Predict the amount of
billings in constant dollars for month 21, using a 99 percent prediction intervaL Interpret
your interval.

```{r}
# get intercept
b0 <- summary(f1)[[4]][1,1]/(1-rho); b0
s_bo <- summary(f1)[[4]][1,2]/(1-rho); s_bo

# get coefficient
b1 <- summary(f1)[[4]][2,1]; b1
s_b1 <- summary(f1)[[4]][2,2]; s_b1

yhat_correct <- b0 + b1*p3$Y
yhat_correct

# Point Forecasts
xn_plus1 <- 3.625
xn <- rev(p3$X)[1]; xn

yhat_nplus1 <- b0+b1*xn_plus1; yhat_nplus1

yn <- rev(p3$Y)[1]; yn
en <- yn - (b0+b1*xn); en

yhat_forecast_nplus1 <- yhat_nplus1 + rho*en
yhat_forecast_nplus1

# Forecast Interval
xprime <- xtnew
xbar_prime <- mean(xprime[-1]) # drop 1st value which is NA
xn_plus1_prime <- xn_plus1 - rho*xn

alpha = 0.05
n <- length(p3$Y)

spred <- sqrt(mse*(1 + (1/n) + (xn_plus1_prime - xbar_prime)^2/(sum((xprime[-1]-xbar_prime)^2))))
spred

pred_l <- yhat_forecast_nplus1 - qt(1-alpha/2, df = n-3)*spred; pred_l
pred_u <- yhat_forecast_nplus1 + qt(1-alpha/2, df = n-3)*spred; pred_u

```



## Problem 4
 
Use question 4 data set, Create development sample (70% of the data) and hold-out sample (30% of the data) use set.seed(1023) before creating the samples.

 a-) Use the development sample , fit a linear regression model,  regression tree and  Neural Network Model, and calculate the SSE for each model, which method has the lowest SSE?
 
```{r}
# load data, examine columns and check for NA, etc.
p4 <- read.csv('Practice Final Question 4.csv')
head(p4)
dim(p4)
str(p4)
summary(p4)

# create train and test set
set.seed(1023)
ind <- sample(1:nrow(p4), round(0.7*nrow(p4)))
tr_p4 <- p4[ind,]
tes_p4 <- p4[-ind,]
#check
dim(tr_p4)

reg4a_lm <- lm(Y~., data = p4)
summary(reg4a_lm)

tree4a <- rpart(Y~., data = tr_p4)
plot_tree4a <- rpart.plot(tree4a, digits = 5, fallen.leaves = TRUE, type = 2, extra = 101)

nn4a <- neuralnet(Y~., data = tr_p4, hidden = c(3,1), act.fct = 'tanh')
nn4a$result.matrix
plot(nn4a, rep = 'best')
```
 
 b-) test the models performances on teh hold out sample, which model would you choose?

```{r}
# lm prediction
predict_4b_lm <- predict(reg4a_lm, tes_p4)
pred_4blm <- data.frame(obs = tes_p4$Y, pred = predict_4b_lm)
defaultSummary(pred_4blm)

# regression tree prediction
predict_4b_tree <- predict(tree4a, tes_p4)
pred_4btree <- data.frame(obs = tes_p4$Y, pred = predict_4b_tree)
defaultSummary(pred_4btree)

# NN prediction
predict_4b_nn <- compute(nn4a, tes_p4)
pred_4bnn <- data.frame(obs = tes_p4$Y, pred = predict_4b_nn)
defaultSummary(pred_4btree)

```

## Problem 5

Use Question 5 dataset, Y is a dichotomous response variable and X2, X3, and X4 are categorical variables. 

a-) Fit a regression model containing the predictor variables in first-order terms and interaction terms for all pairs of predictor variables on development sample. 

```{r}
# load data, examine columns and check for NA, etc.
p5 <- read.csv('Practice Final Question 5.csv')
head(p5)
dim(p5)
str(p5)
summary(p5)

# create train and test set
set.seed(1023)
ind <- sample(1:nrow(p4), round(0.7*nrow(p5)))
tr_p5 <- p5[ind,]
tes_p5 <- p5[-ind,]
#check
dim(tr_p5)

# model with 2 way interactions
reg5a_glm <- glm(Y~.^2, data = tr_p5, family = binomial)
summary(reg5a_glm)

anova(reg5a_glm, test = 'Chi')
```

b-)Use the likelihood ratio test to determine whether all interaction terms can be dropped
from the regression model; State the alternatives, full and reduced models, decision rule, and conclusion. 

```{r}
# reduced model
reg5b_glm <- glm(Y~., data = tr_p5)
summary(reg5b_glm)

# anova test - reduced model, full model
anova(reg5b_glm, reg5a_glm, test = 'Chi')
```

c-)For logistic regression model in part (a), use backward elimination to decide which predictor
variables can be dropped from the regression model.  Which variables are retained in the regression model?

```{r}
drop1(reg5a_glm, test = 'Chi')
```

d-) Conduct the Hosmer-Lemeshow goodness of fit test for the appropriateness of the logistic regression function by forming five groups. State the alternatives, decision rule, and conclusion. 

```{r}
hoslem.test(reg5a_glm$y, fitted(reg5a_glm), g = 5)
# Accept Ho - Fit is good; since p > 0.05

```

e-) Make the prediction for the following two cases and calculate 95% confidence interval

X1=(33,6) X2=(1,1) X3=(1,1) X4=(0,0)

```{r}
test5e<- data.frame(X1=c(33,6), X2=c(1,1), X3=c(1,1), X4=c(0,0)); test5e
predict(reg5a_glm, test5e, interval = 'confidence', level = 0.95)
confint(reg5a_glm) #, level = 0.95)

```

