---
title: "p3_practice_mid_term_hes_e106"
author: "Gurdeep"
date: "21/10/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
rm(list = ls())

setwd("/cloud/project/practice_mid_term") # change to local directory before running

library(knitr)
library(onewaytests)
library(olsrr)
library(lmtest)
library(MASS)
library(ALSM)

pmq3 <- read.csv('PM Q3.csv')
head(pmq3)
colnames(pmq3)
dim(pmq3)

```

Perform one factor analysis by finding the best variable to explain Y. Fit one variable regression model with Y as a dependent variable against remaining variables, as an independent variable one at time. Choice the best variable explain Y and comment on the QQ plot and error vs. fitted values graph for the model assumptions.

```{r}
for (i in 2:(ncol(pmq3))) {
  print(paste("Regression with x",i-1))
  #print(pmq3[1,i])
  reg <- lm(y~pmq3[,i], data = pmq3)
  print(summary(reg))
}
```

As we see from above x2 has the best predictive value. We plot residuals below

```{r}
reg_p3_final <- lm(y~x2, data = pmq3)
summary(reg_p3_final)
ei_p3_final <- reg_p3_final$residuals

# plot residuals and other standard plots
old.par <- par(mfrow = c(2,2))
plot(reg_p3_final)
old.par

newer.par = par(mfrow = c(2,2))
plot(pmq3$x2, pmq3$y, xlab = 'x2', ylab = 'y')
plot(pmq3$x2,ei_p3_final, xlab = 'x', ylab = 'Residuals')
boxplot(ei_p3_final, horizontal = TRUE, staplewax = 0.5, col = 3, xlab = 'Residual')
hist(ei_p3_final, main = 'Histogram of Residuals')
newer.par
```

From above we see that x2 is the best independent variable with R2 of 0.69. It has F* = 85.38 and p-value = 2.926e-11 both of which convey significance. 

The residual vs fitted plots show random pattern which is also reflected in the scatter plot and residual vs X plots.

The QQ plots are inline for 1SD; i.e. 68% of the values and the boxplot is also symmetrical except for a solitary outlier while a few outliers show up in the plots. 

Hence, regression assumptions relating to linear relation, constant variance, independence and normality are satisfied.
