---
title: "p2_practice_mid_term_hes_e106"
author: "Gurdeep"
date: "21/10/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
rm(list = ls())

setwd("/cloud/project/practice_mid_term") # change to local directory before running

library(knitr)
library(onewaytests)
library(olsrr)
library(lmtest)
library(MASS)
library(ALSM)

y=c(98,135,162,178,221,232,283,300,374,395)
x=c(0,1,2,3,4,5,6,7,8,9)

```

2a-) is the linear fit appropriate? If not, transform the data and find an appropriate fit. Comment on the model and regression model assumptions. (20 points)

```{r}
reg2a <- lm(y~x)
summary(reg2a)
ei_2a <- reg2a$residuals

# plot residuals and other standard plots
old.par <- par(mfrow = c(2,2))
plot(reg2a)
old.par

newer.par = par(mfrow = c(2,2))
plot(x, y, xlab = 'x', ylab = 'y')
plot(x,ei_2a, xlab = 'x', ylab = 'Residuals')
boxplot(ei_2a, horizontal = TRUE, staplewax = 0.5, col = 3, xlab = 'Residual')
hist(ei_2a, main = 'Histogram of Residuals')
newer.par

```

At 1st glance from the scatter plot we see what resembles a linear fit. Also we have high R2 (0.9798), high F-test, low p-values for F-test and both predictors, we see from both the residual vs fitted values plot and residuals vs x plots that the regression values have a pattern. 

Hence, since the regression assumptions of normality, independence and correlation are not satisfied we explore transformation of Y only using boxcox transform.

```{r}
boxcox(reg2a, lambda = seq(-1.0, 1.0,0.1))

```

From the boxcox test we see that recommended tranformation is of lambda = 0.5 i.e sqrt(Y). We perform the transformation and explore the results below

```{r}
reg2a_2 <- lm(sqrt(y)~x)
summary(reg2a_2)
ei_2a_2 <- reg2a_2$residuals

# plot residuals and other standard plots
old.par <- par(mfrow = c(2,2))
plot(reg2a_2)
old.par

newer.par = par(mfrow = c(2,2))
plot(x, sqrt(y), xlab = 'x', ylab = 'y')
plot(x,ei_2a_2, xlab = 'x', ylab = 'Residuals')
boxplot(ei_2a_2, horizontal = TRUE, staplewax = 0.5, col = 3, xlab = 'Residual')
hist(ei_2a_2, main = 'Histogram of Residuals')
newer.par

```

From the above plots we see that the regression assumptions are now satisfied. 

2b-) Predict Y when x=10, and calculate the prediction confidence interval for 90% confidence level (10 points)

```{r}

xh <- 10

pred <- predict(reg2a_2, data.frame(x = xh), level = 0.90, interval = 'prediction', se.fit = TRUE )
pred

pred_val <- pred$fit^2
pred_val
```

The final value of Y when x = 10 is 442.0022 with intervals between 408.3675 and 476.9679
